Performing Cluster Analysis on TAD boundaries & Annotations

### 2019-03-22 Scripts/evaluating_pca.Rmd
editted pca script

histonevars <- gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/HistoneModifications"))
tfbsvars <-  gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/tfbs/cell_type_specific/gm12878"))

@@

```{r}
pcadat_gm12878_10kb_count <- pcadat_gm12878_10kb_count[which(rownames(pcadat_gm12878_10kb_count) %in% chromstatevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_count),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_count),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_count)-1)*sum(apply(pcadat_gm12878_10kb_count,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_count,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_count, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_count,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

@@

```{r}
pcadat_gm12878_10kb_perc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_perc.rds")
dim(pcadat_gm12878_10kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_10kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_10kb_perc_f <- pcadat_gm12878_10kb_perc[, -which(colnames(pcadat_gm12878_10kb_perc) %in% nzvar)]

dim(pcadat_gm12878_10kb_perc) 
dim(pcadat_gm12878_10kb_perc_f)

colnames(pcadat_gm12878_10kb_perc_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "perc",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_10kb_perc_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_10kb_perc_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_10kb_perc_f),
          #rowsep=1:nrow(pcadat_gm12878_10kb_perc_f),
          keysize = 1)

#PCA
pcadat_gm12878_10kb_perc_f <- t(pcadat_gm12878_10kb_perc_f)

Group <- ifelse(rownames(pcadat_gm12878_10kb_perc_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_10kb_perc_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_10kb_perc_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_10kb_perc_f <- cbind.data.frame(pcadat_gm12878_10kb_perc_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_f[,-(dim(pcadat_gm12878_10kb_perc_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_10kb_perc_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_f[,-(dim(pcadat_gm12878_10kb_perc_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_10kb_perc_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_perc_f[,-(dim(pcadat_gm12878_10kb_perc_f2)[2])], 4),
         data=pcadat_gm12878_10kb_perc_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

```{r}
pcadat_gm12878_10kb_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_dist.rds")
dim(pcadat_gm12878_10kb_dist)

#performing log2 transformation
pcadat_gm12878_10kb_dist <- apply(pcadat_gm12878_10kb_dist,2,function(x){log(x + 1, base=2)})

#filtering low variance predictors
#nzv <- nearZeroVar(pcadat_gm12878_10kb_dist , saveMetrics= TRUE)
#nzvar <- rownames(nzv[nzv$nzv,])
#nzvar
#no near zero variance predictors

pcadat_gm12878_10kb_dist_f <- pcadat_gm12878_10kb_dist[, -which(colnames(pcadat_gm12878_10kb_dist) %in% nzvar)]
pcadat_gm12878_10kb_dist_f <- pcadat_gm12878_10kb_dist

dim(pcadat_gm12878_10kb_dist) 
dim(pcadat_gm12878_10kb_dist_f)

colnames(pcadat_gm12878_10kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_10kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_10kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_10kb_dist_f),
          #rowsep=1:nrow(pcadat_gm12878_10kb_dist_f),
          keysize = 1)

#PCA
pcadat_gm12878_10kb_dist_f <- t(pcadat_gm12878_10kb_dist_f)

Group <- ifelse(rownames(pcadat_gm12878_10kb_dist_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_10kb_dist_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_10kb_dist_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_10kb_dist_f <- cbind.data.frame(pcadat_gm12878_10kb_dist_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_f[,-(dim(pcadat_gm12878_10kb_dist_f)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_10kb_dist_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_f[,-(dim(pcadat_gm12878_10kb_dist_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_10kb_dist_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_dist_f[,-(dim(pcadat_gm12878_10kb_dist_f2)[2])], 4),
         data=pcadat_gm12878_10kb_dist_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

@@

```{r}
pcadat_gm12878_25kb_count <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/PCA/pcadat_gm12878_25kb_count.rds")
dim(pcadat_gm12878_25kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_25kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_25kb_count_f <- pcadat_gm12878_25kb_count[, -which(colnames(pcadat_gm12878_25kb_count) %in% nzvar)]

dim(pcadat_gm12878_25kb_count) 
dim(pcadat_gm12878_25kb_count_f)

colnames(pcadat_gm12878_25kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_25kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_25kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_25kb_count_f),
          #rowsep=1:nrow(pcadat_gm12878_25kb_count_f),
          keysize = 1)

#PCA
pcadat_gm12878_25kb_count_f <- t(pcadat_gm12878_25kb_count_f)

Group <- ifelse(rownames(pcadat_gm12878_25kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_25kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_25kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_25kb_count_f <- cbind.data.frame(pcadat_gm12878_25kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_count_f[,-(dim(pcadat_gm12878_25kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_25kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_count_f[,-(dim(pcadat_gm12878_25kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_25kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_count_f[,-(dim(pcadat_gm12878_25kb_count_f2)[2])], 4),
         data=pcadat_gm12878_25kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

@@

```{r}
pcadat_gm12878_25kb_perc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/PCA/pcadat_gm12878_25kb_perc.rds")
dim(pcadat_gm12878_25kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_25kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_25kb_perc_f <- pcadat_gm12878_25kb_perc[, -which(colnames(pcadat_gm12878_25kb_perc) %in% nzvar)]

dim(pcadat_gm12878_25kb_perc) 
dim(pcadat_gm12878_25kb_perc_f)

colnames(pcadat_gm12878_25kb_perc_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "perc",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_25kb_perc_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_25kb_perc_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_25kb_perc_f),
          #rowsep=1:nrow(pcadat_gm12878_25kb_perc_f),
          keysize = 1)

#PCA
pcadat_gm12878_25kb_perc_f <- t(pcadat_gm12878_25kb_perc_f)

Group <- ifelse(rownames(pcadat_gm12878_25kb_perc_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_25kb_perc_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_25kb_perc_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_25kb_perc_f <- cbind.data.frame(pcadat_gm12878_25kb_perc_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_perc_f[,-(dim(pcadat_gm12878_25kb_perc_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_25kb_perc_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_perc_f[,-(dim(pcadat_gm12878_25kb_perc_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_25kb_perc_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_perc_f[,-(dim(pcadat_gm12878_25kb_perc_f2)[2])], 4),
         data=pcadat_gm12878_25kb_perc_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

@@

```{r}
pcadat_gm12878_25kb_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/PCA/pcadat_gm12878_25kb_dist.rds")
dim(pcadat_gm12878_25kb_dist)

#performing log2 transformation
pcadat_gm12878_25kb_dist <- apply(pcadat_gm12878_25kb_dist,2,function(x){log(x + 1, base=2)})

#filtering low variance predictors
#nzv <- nearZeroVar(pcadat_gm12878_25kb_dist , saveMetrics= TRUE)
#nzvar <- rownames(nzv[nzv$nzv,])
#nzvar
#no near zero variance predictors

pcadat_gm12878_25kb_dist_f <- pcadat_gm12878_25kb_dist[, -which(colnames(pcadat_gm12878_25kb_dist) %in% nzvar)]
pcadat_gm12878_25kb_dist_f <- pcadat_gm12878_25kb_dist

dim(pcadat_gm12878_25kb_dist) 
dim(pcadat_gm12878_25kb_dist_f)

colnames(pcadat_gm12878_25kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_25kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_25kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_25kb_dist_f),
          #rowsep=1:nrow(pcadat_gm12878_25kb_dist_f),
          keysize = 1)

#PCA
pcadat_gm12878_25kb_dist_f <- t(pcadat_gm12878_25kb_dist_f)

Group <- ifelse(rownames(pcadat_gm12878_25kb_dist_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_25kb_dist_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_25kb_dist_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_25kb_dist_f <- cbind.data.frame(pcadat_gm12878_25kb_dist_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_dist_f[,-(dim(pcadat_gm12878_25kb_dist_f)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_25kb_dist_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_dist_f[,-(dim(pcadat_gm12878_25kb_dist_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_25kb_dist_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_dist_f[,-(dim(pcadat_gm12878_25kb_dist_f2)[2])], 4),
         data=pcadat_gm12878_25kb_dist_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

```{r}
pcadat_gm12878_50kb_count <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/PCA/pcadat_gm12878_50kb_count.rds")
dim(pcadat_gm12878_50kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_50kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_50kb_count_f <- pcadat_gm12878_50kb_count[, -which(colnames(pcadat_gm12878_50kb_count) %in% nzvar)]

dim(pcadat_gm12878_50kb_count) 
dim(pcadat_gm12878_50kb_count_f)

colnames(pcadat_gm12878_50kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_50kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_50kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_50kb_count_f),
          #rowsep=1:nrow(pcadat_gm12878_50kb_count_f),
          keysize = 1)

#PCA
pcadat_gm12878_50kb_count_f <- t(pcadat_gm12878_50kb_count_f)

Group <- ifelse(rownames(pcadat_gm12878_50kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_50kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_50kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_50kb_count_f <- cbind.data.frame(pcadat_gm12878_50kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_50kb_count_f[,-(dim(pcadat_gm12878_50kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_50kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_50kb_count_f[,-(dim(pcadat_gm12878_50kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_50kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_50kb_count_f[,-(dim(pcadat_gm12878_50kb_count_f2)[2])], 4),
         data=pcadat_gm12878_50kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

@@

```{r}
pcadat_gm12878_50kb_perc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/PCA/pcadat_gm12878_50kb_perc.rds")
dim(pcadat_gm12878_50kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_50kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_50kb_perc_f <- pcadat_gm12878_50kb_perc[, -which(colnames(pcadat_gm12878_50kb_perc) %in% nzvar)]

dim(pcadat_gm12878_50kb_perc) 
dim(pcadat_gm12878_50kb_perc_f)

>>>>

tfbsvars <-  gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/tfbs/cell_type_specific/gm12878"))

@@

```{r}
pcadat_gm12878_10kb_count_chromstates <- pcadat_gm12878_10kb_count[which(rownames(pcadat_gm12878_10kb_count) %in% chromstatevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_chromstates),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_chromstates),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_count_chromstates)-1)*sum(apply(pcadat_gm12878_10kb_count_chromstates,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_count_chromstates,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_count_chromstates, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_count_chromstates,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

@@

##### Overlap Percents

```{r}
pcadat_gm12878_10kb_perc_chromstates <- pcadat_gm12878_10kb_perc[which(rownames(pcadat_gm12878_10kb_perc) %in% chromstatevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_chromstates),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_chromstates),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_perc_chromstates)-1)*sum(apply(pcadat_gm12878_10kb_perc_chromstates,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_perc_chromstates,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_perc_chromstates, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_perc_chromstates,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

@@

#####Distance

```{r}
pcadat_gm12878_10kb_dist_chromstates <- pcadat_gm12878_10kb_dist[which(rownames(pcadat_gm12878_10kb_dist) %in% chromstatevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_chromstates),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_chromstates),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_dist_chromstates)-1)*sum(apply(pcadat_gm12878_10kb_dist_chromstates,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_dist_chromstates,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_dist_chromstates, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_dist_chromstates,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

@@

#### Histone Modifications

```{r}
histonevars <- gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/HistoneModifications"))
```

```{r}
pcadat_gm12878_10kb_count_histones <- pcadat_gm12878_10kb_count[which(rownames(pcadat_gm12878_10kb_count) %in% histonevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_histones),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_histones),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_count_histones)-1)*sum(apply(pcadat_gm12878_10kb_count_histones,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_count_histones,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_count_histones, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_count_histones,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

@@

##### Overlap Percents

```{r}
pcadat_gm12878_10kb_perc_histones <- pcadat_gm12878_10kb_perc[which(rownames(pcadat_gm12878_10kb_perc) %in% histonevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_histones),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_histones),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_perc_histones)-1)*sum(apply(pcadat_gm12878_10kb_perc_histones,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_perc_histones,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_perc_histones, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_perc_histones,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

@@

#####Distance

```{r}
pcadat_gm12878_10kb_dist_histones <- pcadat_gm12878_10kb_dist[which(rownames(pcadat_gm12878_10kb_dist) %in% histonevars),]
